A comparison of different methods for calculating tangent-stiffness matrices in a massively parallel computational peridynamics code.

### 2013-10-25 jacobian_calc_CMAME.tex
figure aspect and axes
More incremental fixes

In detail, a tangent-stiffness matrix can be described as a collection of the first-order partial derivatives of a vector valued function with respect to each independent vector component of the function for a given vector.   A general mathematical formula for a tangent-stiffness matrix, $K_{ij}$,
can be expressed in indicial notation as 
%
\begin{equation} 
  K_{ij} = \left. \frac{\partial F_i}{\partial X_j}\right|_{\vec{X}_0}
\end{equation}
%
Where $F_i$ is the $i^{\mbox{th}}$ component of the vector valued function, $X_j$ is the $j^{\mbox{th}}$ component of the vector argument to $F$, and a particular value of the vector is chosen to linearize about $\vec{X}_0$. One then evaluates the expression for each combination $i, j$ corresponding to an (row, column) location in the tangent-stiffness matrix. By inspection, the elements of a tangent-stiffness matrix can be estimated using the any of the complex-step, AD, or common finite-difference techniques for functions $F:R^1 \rightarrow R^1$, since taking partial derivatives entails holding all but a single independent vector component of the function constant, and each element of the function can be evaluated independently of the others. \\

\subsubsection{Differencing formulas for tangent-stiffness matrix evaluation with each method.}
For reference, expressions for calculating tangent stiffness matrices with CTSE, AD,
central-difference and forward-difference will be shown and explained. The expressions for AD,
central-difference and forward-difference are taken from \cite{ref-Adaggio}, while the expression
for CTSE is original.

When the using the forward-difference technique the formula for calculating a tangent-stiffness
matrix is:

\begin{equation} K_{ij} = \frac{\partial F_i^{int}(x + \delta e_j) - F_i^{int}(x)}{\delta}
\end{equation}

Where $K_{ij}$ is the index notation representation of an element of the tangent-stiffness matrix at
row $i$, column $j$. The expression says that the partial derivative of the $i$'th component of $F$
w.r.t the $j$'th component of $x$ is equal to the quotient of the partial difference of two terms
and a denomenator. The first of the 'two terms' is $F_i$ evaluated at $x$ where
the $j$'th component of $x$ has been perturbed by an very small amount in the direction of the
positive $x_j$ axis. The second of the 'two terms' is $F_I$ evaluated at an unperturbed $x$. The
denominator is the magnitude of perturbation called "probe distance" \cite{ref-Adaggio}

>>>>

In detail, a tangent-stiffness matrix can be described as a collection of the first-order partial derivatives of a vector valued function with respect to each independent vector component of the function for a given vector.   A general mathematical formula for a tangent-stiffness matrix, $K_{ij}$,
can be expressed in indicial notation as 
%
\begin{equation} 
  K_{ij} = \left. \frac{\partial F_i}{\partial X_j}\right|_{\vec{X}_0}
\end{equation}
%
Where $F_i$ is the $i^{\mbox{th}}$ component of the vector valued function, $X_j$ is the $j^{\mbox{th}}$ component of the vector argument to $F$, and a particular value of the vector is chosen to linearize about $\vec{X}_0$. One then evaluates the expression for each combination $i, j$ corresponding to a (row, column) location in the tangent-stiffness matrix. By inspection, the elements of a tangent-stiffness matrix can be estimated using the any of the complex-step, AD, or common finite-difference techniques for functions $F:R^1 \rightarrow R^1$, since taking partial derivatives entails holding all but a single independent vector component of the function constant, and each element of the function can be evaluated independently of the others. \\

\subsubsection{Differencing formulas for tangent-stiffness matrix evaluation.}

For reference, expressions for calculating tangent stiffness matrices with complex-step, automatic differentiation, central-difference and forward-difference will be presented and explained.

When the using the forward-difference technique the formula for calculating a tangent-stiffness
matrix is
%
\begin{equation} 
  K_{ij} = \frac{\partial F_i^{int}(x + h e_j) - F_i^{int}(x)}{h}
\end{equation}
%
Where $K_{ij}$ is the indicial notation representation of an element of the tangent-stiffness matrix at row $i$, column $j$. The expression says that the partial derivative of the $i$'th component of $F$ with respect the $j$'th component of $x$ is equal to the quotient of the partial difference of two terms
and a denomenator. The first of the 'two terms' is $F_i$ evaluated at $x$ where
the $j$'th component of $x$ has been perturbed by an very small amount in the direction of the
positive $x_j$ axis. The second of the 'two terms' is $F_I$ evaluated at an unperturbed $x$. The
denominator is the magnitude of perturbation called "probe distance" \cite{ref-Adaggio}
